---
title: "Cox Proportional Hazards analysis"
output: html_notebook
---

# Load data
Load clinical data and append gene expression data to the dataframe
```{r}
library(dplyr)
library(tibble)
library(tidyverse)

# genes in selected pathway
geneset <- c(Gene1, Gene2, Gene3...)

# gene expression data
expdata <- read.table("microarray/GSEnnnnn",sep=",",header=T) # or read.delim
rownames(expdata) <- expdata$Gene
# filter selected genes
expdata <- expdata %>%
  filter(row.names(expdata) %in% geneset)

# clinic data
clindata <- read.table("clinical/GSEnnnnn",sep=",",header=T)
rownames(clindata) <- clindata$Sample

# append exp data to clinical data
get_exp <- function(x,gene){ expdata[gene,x] }
for (gene in geneset){
  clindata[gene] <- sapply(clindata$Sample,get_exp,gene=gene)
}
```

# Cox Proportional Hazards (CPH) Model: univariate
Install required libraries and call the CPH regression model in R
```{r}
## if the packages are not installed:
# install.packages(c("survival","survminer"))
library(survival)
library(survminer)
clindata$gene_quartile <- ntile(clindata[,Gene1], n=4)
temp <- clindata[, c("OS_time","OS_status",Gene1)]
# to filter out a specific subset of the cohort for the CPH model
# clindata <- clindata[clindata$gene_quartile==2,] 
res.cox <- coxph(Surv(as.numeric(OS_time),as.numeric(OS_status)) ~ Gene1, data=temp)
summary(res.cox)
```

# Notes
1. “z” gives the Wald statistic value
2. Sign of the regression coefficients (coef): A positive sign means that the hazard (risk of death) is higher, 
   and thus the prognosis worse, for subjects with higher values of that variable
